### 2014-05-09 latex/paper1.tex
starting paper1

\documentclass[12pt, oneside, letterpaper, fleqn]{article}

\usepackage{bobianite}

\pagestyle{fancy}
\headheight 1.6cm
\fancyhead[L]{Bryance Oyang\\HPS 160}

\begin{document}
\begin{center}
\textbf{title}
\end{center}

\nocite{*}
\bibliographystyle{unsrt}
\bibliography{bib_paper1}

\end{document}

### 2014-05-09 latex/outline_paper1.tex
wrote zeeman stuff

\documentclass[12pt, oneside, letterpaper, fleqn]{article}

\usepackage{bobianite}

\begin{document}
\begin{center}
\textbf{Outline}
\end{center}
\begin{itemize}

\item physics and technology for experiment had progressed enough that
it was possible for the first time to investigate the structure of
matter. physics evolved from being concerned with the behavior of things
to what makes up those things, what are the constituents of matter,
energy, light, sound.

\item Zeeman
\begin{itemize}
\item a good amount of Zeeman's paper was devoted to investigating the
structure of atoms
\item he put heavy emphasis on what his experimental results implied
about the structure of matter
\item zeeman discusses his correspondence with Lorentz: a model for the
atom and how the model was in agreement with experiment
\item with the model he estimated the charge to mass ratio of the
charged particles in the atom
\item faraday had previously tried this, but did not have precise
experimental apparatus that zeeman had
\item zeeman performed several versions of the experiment: in both he
observed a widening of the spectral line of atoms when a magnetic field
was applied
\item lorentz and zeeman's model of the atom predicted the line would
split into 3 lines. the 3 lines blurred together would look as if the
line widened
\end{itemize}

\end{itemize}
\end{document}

### 2014-05-09 latex/paper1.tex
wrote zeeman stuff

@@
 \usepackage{bobianite}
+\setstretch{2}
 
@@
 
+Around 1896--1897, Pieter Zeeman investigated the effects that a
+magnetic field would have on the spectral lines of atoms. Michael
+Faraday had previously investigated this same problem, but at the time
+he lacked the technology and equipment to make the precise measurements
+that were needed, and as a result did not find anything interesting.
+Zeeman, on the other hand, was able to discover what is now called the Zeeman
+effect. He vaporized some sodium in a small container and shone a light
+through it. The spectrum of the light had the dark absorption lines of
+sodium. Then Zeeman applied a magnetic field to the sodium container,
+and he observed that the absorption lines instantly widened. After
+corresponding with Lorentz, Zeeman published his results.
+In Zeeman's publication, he does not simply describe his
+observations of the effects the magnetic field had on the spectral lines
+though. He puts a heavy emphasis on what his
+experimental results implied about the structure of matter and boldly hypothesizes
+a model of the atom, with help from Lorentz.
+
+In this model, charged particles orbit a central nucleus in circles.
+When a magnetic field is applied, the orbits are perturbed, so that the
+frequency of light emitted or absorbed by the atom changes slightly. The
+result is that one spectral line splits into three, and therefore looks
+as if the line widened. Using Maxwell's electromagnetic theory and
+assuming this model of the atom, Lorentz
+predicted the polarization of light for each of the three split lines,
+and his prediction was found to be in exact agreement with Zeeman's
+experiment. All this was done even before the discovery of the electron by JJ
+Thompson, showing that there was an intense interest and rush to
+understand the structure of matter.
+
 \nocite{*}

### 2014-05-15 latex/paper1.tex
almost done, just need intro and conclusion

@@
 \pagestyle{fancy}
-\headheight 1.6cm
 \fancyhead[L]{Bryance Oyang\\HPS 160}
@@
 \begin{center}
-\textbf{title}
+\textbf{Transformative discoveries in physics at the turn of the 20th
+century}
 \end{center}
 
-Around 1896--1897, Pieter Zeeman investigated the effects that a
-magnetic field would have on the spectral lines of atoms. Michael
-Faraday had previously investigated this same problem, but at the time
-he lacked the technology and equipment to make the precise measurements
-that were needed, and as a result did not find anything interesting.
-Zeeman, on the other hand, was able to discover what is now called the Zeeman
-effect. He vaporized some sodium in a small container and shone a light
-through it. The spectrum of the light had the dark absorption lines of
-sodium. Then Zeeman applied a magnetic field to the sodium container,
-and he observed that the absorption lines instantly widened. After
-corresponding with Lorentz, Zeeman published his results.
-In Zeeman's publication, he does not simply describe his
-observations of the effects the magnetic field had on the spectral lines
-though. He puts a heavy emphasis on what his
-experimental results implied about the structure of matter and boldly hypothesizes
-a model of the atom, with help from Lorentz.
+Wilhelm R\"ontgen, in 1895, discovered X-rays, a new kind of ray that
+penetrates many substances including human flesh. Inspired by
+R\"ontgen's discovery, Henri Becquerel and Marie and Pierre Curie began
+investigations in radioactive substances and the various rays those
+substances produced. They found that radioactive substances would
+spontaneously emit energetic rays and would not require an outside
+source of energy to do this. These discoveries made physicists wonder
+what was it about the structure of matter that allowed these rays to be
+created, and also created interest in the composition of the emitted
+rays.
+
+Further investigations revealed that there were three main types of rays
+emitted by radioactive substances, now called $\alpha$, $\beta$, and
+$\gamma$ rays. The $\alpha$ and $\beta$ rays were able to be deflected
+by magnetic fields, and physicists therefore deduced they were charged.
+On the other hand, $\gamma$ rays were not deflected by magnetic fields
+and behaved similarly with X-rays. Physicists hypothesized that the
+emission of these rays should be associated with a ``slow modification
+of the atoms of the radioactive substances'' \cite{becquerel}, but were
+unsure whether the modification was brought about by additional
+invisible rays or was simply spontaneous. These experiments in
+radioactivity and the hypotheses developed around them demonstrates that
+at the turn of the 20th century, a great interest in the structure of
+matter and atomic theory was emerging.
+
+In 1897, J.\ J.\ Thomson conducted a series of experiments to study
+cathode ray tubes. There was a disagreement among English and German
+physicists at the time whether these cathode rays were material
+(supported by English physicists) or ethereal (supported by German
+physicists) in nature. Thomson showed that the cathode rays were
+deflected by both electric and magnetic fields and that they were
+negatively charged. He also determined that the rays were composed of
+material particles and found the charge-to-mass ratio of these
+particles, now called electrons. Thomson noticed that the material of
+the cathode in the tube had no effect on the charge-to-mass ratio of the
+particles in the rays, suggesting that all cathode rays were composed of
+the same kind of particle.
+
+Thomson's discovery of what made up cathode rays was considered very
+important at the time, and he was awarded the Nobel Prize in Physics in
+1906 for his efforts. The discovery of the electron also helped to fuel
+futher interest in the structure of matter. Thomson later proposed the
+plum pudding model of the atom which had electrons embedded in a sea of
+positive charge.
+
+Around the same time that Thomson was investigating cathode rays, Pieter
+Zeeman investigated the effects that a magnetic field would have on the
+spectral lines of atoms. Michael Faraday had previously investigated a
+similar problem, but at the time he lacked the technology and equipment
+to make the precise measurements that were needed, and as a result did
+not find anything interesting.  Zeeman, on the other hand, was able to
+discover what is now called the Zeeman effect. He vaporized some sodium
+in a small container and shone a light through it. The spectrum of the
+light had the dark absorption lines of sodium. Then Zeeman applied a
+magnetic field to the sodium container, and he observed that the
+absorption lines instantly widened.  After corresponding with Lorentz,
+Zeeman published his results.  In Zeeman's publication, he does not
+simply describe his observations of the effects the magnetic field had
+on the spectral lines though. He puts a heavy emphasis on what his
+experimental results implied about the structure of matter and boldly
+hypothesizes a model of the atom, with help from Lorentz.
 
@@
 as if the line widened. Using Maxwell's electromagnetic theory and
-assuming this model of the atom, Lorentz
-predicted the polarization of light for each of the three split lines,
-and his prediction was found to be in exact agreement with Zeeman's
-experiment. All this was done even before the discovery of the electron by JJ
-Thompson, showing that there was an intense interest and rush to
-understand the structure of matter.
+assuming this model of the atom, Lorentz predicted the polarization of
+light for each of the three split lines, and his prediction was found to
+be in exact agreement with Zeeman's experiment. This was done around the
+time, if not before the discovery of the electron by J.\ J.\ Thomson,
+showing that there was an intense interest and rush to understand the
+structure of matter. Furthermore, this model of the atom along with the
+experiment allowed Zeeman to calculate the charge-to-mass ratio of the
+electron, which was in agreement with the value obtained by J.\ J.\
+Thomson.
+
+Besides an intense interest in the structure of matter at the turn of
+the century, Max Planck's explanation of blackbody radiation and
+Einstein's subsequent explanation of the photoelectric effect created an
+interest in the structure of light. After Maxwell formulated his Maxwell
+equations for electromagnetic field, it was realized that light is
+electromagnetic radiation. However, when physicists applied Maxwell's
+theory and statistical mechanics to try to understand the spectrum of
+blackbody radiation, they found a large disagreement between theory and
+experiment. Max Planck proposed a fix to the situation, where he
+hypothesized that a blackbody could only emit discrete packets of energy
+as light, and the size of the packets depended on the frequency of light
+being emitted. A few years later, Einstein explained the photoelectric
+effect by hypothesizing that light was composed of particles called
+photons, and the energy of these photons was connected to the frequency
+of light in the same way as in Planck's theory.
 
 \nocite{*}
-\bibliographystyle{unsrt}
+%\bibliographystyle{unsrt}
+\bibliographystyle{plain}
 \bibliography{bib_paper1}

### 2014-05-15 latex/paper1.tex
i think i'm done

@@
 \end{center}
+
+At the turn of the 20th century, physics had progressed to the point
+where it was now possible to investigate the structure of matter and
+light. Electromagnetic theory had been developed so that physicists
+could investigate the charged particles making up the atom. Statistical
+mechanics, originally developed to explain heat and gases, would help
+Planck and Einstein realize the corpuscular nature of light. Physics was
+ready for a major transformation. In the past, physics had been
+primarily concerned with the behavior and motion of things, such as the
+motion of planets, the behavior of light, the behavior of electric and
+magnetic fields, or the behavior of heat and gases. In this paper, I
+argue that at the turn of the 20th century, physics shifted from being
+mainly concerned with the behavior of matter and light to being
+interested in the actual structure of matter and light. Physicists
+became interested in what made up the the things they were studying,
+what was inside the atom, what was light made of, and physicists finally
+had the tools they needed to find answers to these questions. I will
+use the investigations into radioactivity, Thomson's discovery of the
+electron, Zeeman's experiments, and Planck and Einstein's ideas about
+light to illustrate the heightened interest in the structure of matter
+of light during this time period.
 
@@
 experiment allowed Zeeman to calculate the charge-to-mass ratio of the
-electron, which was in agreement with the value obtained by J.\ J.\
-Thomson.
+electron, which was in agreement with the value obtained more directly
+by J.\ J.\ Thomson.
 
@@
 
+These experiments and investigations during the turn of the century
+propelled physics into a new era. The intense interest in the atom and
+the nature of light created during this time fueled the quantum
+revolution taking place twenty years later. Throughout the 20th century,
+a large part of physics continued to investigate the structure of matter
+and light. Many new particles and families of particles were discovered,
+such as neutrinos, quarks, force carriers, and very recently the Higgs
+boson. An important breakthrough during the turn of the 20th century
+that I did not mention earlier in this paper is Einstein's theory of
+relativity. It was one of the few accomplishments during this time that
+was not concerned with the structure of matter or light, but was
+concerned with the nature of space and time. This theory later lead to
+Einstein's theory of gravitation. Interestingly, while all aspects
+of physics that contributed to understanding the structure of matter and
+light at the time have now been reformulated as quantum theories, such
+as electromagnetic theory or statistical mechanics, Einstein's theory of
+gravity, which did not come from investigating the structure of matter
+or light, to this day has no satisfactory quantum counterpart.
+
 \nocite{*}

### 2014-05-15 latex/paper1.tex
minor spellcheck/other changes

@@
 mainly concerned with the behavior of matter and light to being
-interested in the actual structure of matter and light. Physicists
-became interested in what made up the the things they were studying,
-what was inside the atom, what was light made of, and physicists finally
-had the tools they needed to find answers to these questions. I will
-use the investigations into radioactivity, Thomson's discovery of the
-electron, Zeeman's experiments, and Planck and Einstein's ideas about
-light to illustrate the heightened interest in the structure of matter
-of light during this time period.
+interested in the actual \textit{structure} of matter and light.
+Physicists became interested in what made up the the things they were
+studying, what was inside the atom, what was light made of, and
+physicists finally had the tools they needed to find answers to these
+questions. I will use the investigations into radioactivity, Thomson's
+discovery of the electron, Zeeman's experiments, and Planck and
+Einstein's ideas about light to illustrate the heightened interest in
+the structure of matter of light during this time period.
 
@@
 R\"ontgen's discovery, Henri Becquerel and Marie and Pierre Curie began
-investigations in radioactive substances and the various rays those
+investigations into radioactive substances and the various rays those
 substances produced. They found that radioactive substances would
@@
 1906 for his efforts. The discovery of the electron also helped to fuel
-futher interest in the structure of matter. Thomson later proposed the
+further interest in the structure of matter. Thomson later proposed the
 plum pudding model of the atom which had electrons embedded in a sea of

### 2014-06-07 latex/paper2.tex
starting final paper

--- 2014-05-15 latex/paper1.tex
@@
 \begin{center}
-\textbf{Transformative discoveries in physics at the turn of the 20th
-century}
+\textbf{}
 \end{center}
-
-Wilhelm R\"ontgen, in 1895, discovered X-rays, a new kind of ray that
-penetrates many substances including human flesh. Inspired by
-R\"ontgen's discovery, Henri Becquerel and Marie and Pierre Curie began
-investigations in radioactive substances and the various rays those
-substances produced. They found that radioactive substances would
-spontaneously emit energetic rays and would not require an outside
-source of energy to do this. These discoveries made physicists wonder
-what was it about the structure of matter that allowed these rays to be
-created, and also created interest in the composition of the emitted
-rays.
-
-Further investigations revealed that there were three main types of rays
-emitted by radioactive substances, now called $\alpha$, $\beta$, and
-$\gamma$ rays. The $\alpha$ and $\beta$ rays were able to be deflected
-by magnetic fields, and physicists therefore deduced they were charged.
-On the other hand, $\gamma$ rays were not deflected by magnetic fields
-and behaved similarly with X-rays. Physicists hypothesized that the
-emission of these rays should be associated with a ``slow modification
-of the atoms of the radioactive substances'' \cite{becquerel}, but were
-unsure whether the modification was brought about by additional
-invisible rays or was simply spontaneous. These experiments in
-radioactivity and the hypotheses developed around them demonstrates that
-at the turn of the 20th century, a great interest in the structure of
-matter and atomic theory was emerging.
-
-In 1897, J.\ J.\ Thomson conducted a series of experiments to study
-cathode ray tubes. There was a disagreement among English and German
-physicists at the time whether these cathode rays were material
-(supported by English physicists) or ethereal (supported by German
-physicists) in nature. Thomson showed that the cathode rays were
-deflected by both electric and magnetic fields and that they were
-negatively charged. He also determined that the rays were composed of
-material particles and found the charge-to-mass ratio of these
-particles, now called electrons. Thomson noticed that the material of
-the cathode in the tube had no effect on the charge-to-mass ratio of the
-particles in the rays, suggesting that all cathode rays were composed of
-the same kind of particle.
-
-Thomson's discovery of what made up cathode rays was considered very
-important at the time, and he was awarded the Nobel Prize in Physics in
-1906 for his efforts. The discovery of the electron also helped to fuel
-futher interest in the structure of matter. Thomson later proposed the
-plum pudding model of the atom which had electrons embedded in a sea of
-positive charge.
-
-Around the same time that Thomson was investigating cathode rays, Pieter
-Zeeman investigated the effects that a magnetic field would have on the
-spectral lines of atoms. Michael Faraday had previously investigated a
-similar problem, but at the time he lacked the technology and equipment
-to make the precise measurements that were needed, and as a result did
-not find anything interesting.  Zeeman, on the other hand, was able to
-discover what is now called the Zeeman effect. He vaporized some sodium
-in a small container and shone a light through it. The spectrum of the
-light had the dark absorption lines of sodium. Then Zeeman applied a
-magnetic field to the sodium container, and he observed that the
-absorption lines instantly widened.  After corresponding with Lorentz,
-Zeeman published his results.  In Zeeman's publication, he does not
-simply describe his observations of the effects the magnetic field had
-on the spectral lines though. He puts a heavy emphasis on what his
-experimental results implied about the structure of matter and boldly
-hypothesizes a model of the atom, with help from Lorentz.
-
-In this model, charged particles orbit a central nucleus in circles.
-When a magnetic field is applied, the orbits are perturbed, so that the
-frequency of light emitted or absorbed by the atom changes slightly. The
-result is that one spectral line splits into three, and therefore looks
-as if the line widened. Using Maxwell's electromagnetic theory and
-assuming this model of the atom, Lorentz predicted the polarization of
-light for each of the three split lines, and his prediction was found to
-be in exact agreement with Zeeman's experiment. This was done around the
-time, if not before the discovery of the electron by J.\ J.\ Thomson,
-showing that there was an intense interest and rush to understand the
-structure of matter. Furthermore, this model of the atom along with the
-experiment allowed Zeeman to calculate the charge-to-mass ratio of the
-electron, which was in agreement with the value obtained by J.\ J.\
-Thomson.
-
-Besides an intense interest in the structure of matter at the turn of
-the century, Max Planck's explanation of blackbody radiation and
-Einstein's subsequent explanation of the photoelectric effect created an
-interest in the structure of light. After Maxwell formulated his Maxwell
-equations for electromagnetic field, it was realized that light is
-electromagnetic radiation. However, when physicists applied Maxwell's
-theory and statistical mechanics to try to understand the spectrum of
-blackbody radiation, they found a large disagreement between theory and
-experiment. Max Planck proposed a fix to the situation, where he
-hypothesized that a blackbody could only emit discrete packets of energy
-as light, and the size of the packets depended on the frequency of light
-being emitted. A few years later, Einstein explained the photoelectric
-effect by hypothesizing that light was composed of particles called
-photons, and the energy of these photons was connected to the frequency
-of light in the same way as in Planck's theory.
 
@@
 \bibliographystyle{plain}
-\bibliography{bib_paper1}
+\bibliography{bib_paper2}
 

### 2014-06-08 latex/paper2.tex
did bibliography
wrote 4 pages

@@
 \begin{center}
-\textbf{}
+\textbf{P. A. M. Dirac and His Search for Beautiful Mathematics}
 \end{center}
+
+Dirac's style and approach to his work has had an enormous influence
+on theoretical physics. To fully understand Dirac's style of thinking,
+we must begin by looking at Dirac's education and the events that
+influenced him.
+
+Dirac began studying engineering at the University of Bristol when he
+was 16. During his time at the university, he read in \textit{The Times}
+about the verification of Einstein's general theory of relativity, which
+had a huge influence on him and sparked his interest in physics. Eager
+to find out more about Einstein's theory, Dirac was disappointed when he
+realized that none of his teachers knew much about relativity either.
+Fortunately for Dirac, the philosopher Charlie Broad, who had studied the
+Einstein's theory, was appointed as the Professor of Philosphy at
+Bristol in Dirac's final year, and gave lectures on relativity. From
+these lectures and from reading Arthur Eddington's book on relativity,
+Dirac was able to learn both the special and general theory of
+relativity.
+
+While Einstein's theories sparked Dirac's interest in physics, Dirac's
+education in engineering helped him develop an appreciation for
+approximate theories, because engineers had to constantly make
+approximations. Dirac realized that Newtonian mechanics can be viewed as
+an approximation to Einstein's more refined relativistic mechanics, and
+perhaps Einstein's theory is also an approximation to reality. Dirac
+developed the belief that the purpose of science is to develop ever more
+accurate approximations, and that an exact theory might never be found.
+Dirac thought that ``even approximate theories can have mathematical
+beauty'' \cite[pg. 45]{strangest_man} as a result of his engineering training,
+and this point of view of science as approximations has influenced
+Richard Feynman and many other physicists today.
+
+Despite graduating with first class honors, Dirac was unable to find a
+job after he completed his engineering degree. Recognizing Dirac's
+mathematical talent, the faculty at the University of Bristol convinced
+him to return to school to study mathematics. While studying mathematics
+Dirac learned of the work of William Hamilton, who had reformulated
+Newton's laws using a new approach, called Hamiltonian mechanics. A few
+years later, Dirac would use the ideas of Hamiltonian mechanics to set up a
+logical framework for quantum mechanics.
+
+After completing his mathematics degree, again with first class honors,
+Dirac got a place at Cambridge to continue his studies. Dirac's passion
+was still in the theory of relativity, but Ralph Fowler was assigned to
+be his supervisor. Fowler was a physicist who was an expert in atomic
+theory but, disappointingly for Dirac, was not an expert in relativity.
+Fowler introduced Dirac to Bohr's model of the atom and the emerging
+ideas of quantum theory at the time, but Dirac continued to pursue his
+interest in relativity. He attended the lectures of Arthur Eddington,
+where he ``admired Eddington's mathematical approach to science, which
+would become one of the most powerful influences on him'' \cite[pg.
+61]{strangest_man}. This style was in contrast to that of Ernest
+Rutherford, another leading physicist at Cambridge, who disliked
+theorising and is noted for saying ``Don't let me catch anyone talking
+about the universe in my department'' \cite[pg. 61]{strangest_man}.
+
+Dirac was a hard-working student at Cambridge. He worked almost all day
+for six days every week, reading books, journals, and reviewing his
+lecture notes, only taking time off on Sundays. On Sundays, he would
+take long walks that lasted the whole day, and tried not to think about
+his work so that he could return to it fresh on Mondays. Dirac's
+supervisor, Fowler, introduced him to atomic theory and Dirac spent many
+of his days reading Arnold Sommerfeld's book on atomic theory.
+Sommerfeld improved on the Bohr model of the atom by using Einstein's
+relativistic mechanics instead of Newtonian mechanics to describe the
+orbit of electrons around the nucleus, and Sommerfeld was able to obtain
+more accurate calculations of the energy levels of the atom compared to
+Bohr's original model. Dirac, being interested in relativity, must have
+been influenced greatly by this work, since years later he would
+merge relativity and quantum mechanics.
+
+In September 1925, Fowler received a paper from Heisenberg, and Fowler
+promptly forwarded it to Dirac. Heisenberg presented a new view of the
+atom. Heisenberg felt that the Bohr model, which talked about the
+position of the electron and the time it takes for it to orbit the
+nucleus, must be wrong, because these quantities could not actually be
+measured in experiment. Heisenberg felt that the theory should only
+establish relations between quantities that could be measured, and
+developed an abstract but working theory using matrices. Heisenberg's
+philosophy of only talking about things that could be measured had an
+influence on Dirac, and Dirac would later write in his famous book
+\textit{The Principles of Quantum Mechanics} \cite{principles} that
+questions about things that cannot be investigated by experiment should
+be considered as outside the domain of science, a view that has
+influenced Feynman as can be seen from Feynman's lectures.
+
+What was strange to Heisenberg was that the order in which matrices are
+multiplied matters, $a$ times $b$ could be different from $b$ times $a$.
+This $ab - ba$ caught Dirac's attention, as Dirac was familiar with this
+noncommuting multiplication from his mathematics training, but did not
+understand its physical significance or implications. Always interested
+in relativity, Dirac also tried to create a relativistic version of
+Heisenberg's theory but had no success and soon gave up.
+
+On one of Dirac's relaxing Sunday walks, it occured to him that the
+noncommuting matrices in Heisenberg's theory might be related to the
+Poisson bracket, a concept he encountered while studying Hamilton's
+theory in classical mechanics as a mathematics student a few years ago.
+However, Dirac did not remember exactly what a Poisson bracket was at
+the time, and all the libraries were closed on Sunday. After waiting
+impatiently and eagerly through the night, Dirac went to the libraries
+on Monday morning to look up the Poisson bracket. After a few more weeks
+of work, Dirac was able to complete the connection of the $ab - ba$ in
+Heisenberg's work with the Poisson bracket in classical mechanics, and
+realized he could build a consistent quantum mechanics through this
+correspondence with classical mechanics. Dirac's supervisor Fowler
+realized the importance of Dirac's work and urged the Royal Society to
+quickly publish Dirac's paper. Dirac arrived at these ideas through
+mathematical reasoning and analogy, as opposed to physical intuition,
+showing the influence of Dirac's mathematical training. A few years
+later, Dirac would return to this idea to develop quantum
+electrodynamics by making the noncommutivity of quantum fields
+correspond to classical Poisson brackets.
+
+Despite using mathematics as a primary tool for his thinking, Dirac did
+mathematics quite differently from mathematicians. He says ``I think you
+can see here the effects of an engineering training. I just wanted to
+get results quickly, results which I felt one could have some confidence
+in, even though they did not follow from strict logic, and I was using
+the mathematics of engineers, rather than the rigorous mathematics which
+had been taught to me by Fraser'' \cite[pg. 95]{strangest_man}. Dirac's
+style and approach to mathematical physics continues even today in
+physics, as most physicists avoid the level of rigor that mathematicians
+typically employ.
 

### 2014-06-08 latex/paper2.tex
have 6 pages

@@
 \begin{center}
-\textbf{P. A. M. Dirac and His Search for Beautiful Mathematics}
+\textbf{P. A. M. Dirac and his Search for Mathematical Beauty in Nature}
 \end{center}
@@
 Dirac began studying engineering at the University of Bristol when he
-was 16. During his time at the university, he read in \textit{The Times}
-about the verification of Einstein's general theory of relativity, which
-had a huge influence on him and sparked his interest in physics. Eager
-to find out more about Einstein's theory, Dirac was disappointed when he
+was 16. During his time at the university, he read in a newspaper about
+the verification of Einstein's general theory of relativity, which had a
+huge influence on him and sparked his interest in physics. Eager to find
+out more about Einstein's theory, Dirac was disappointed when he
 realized that none of his teachers knew much about relativity either.
-Fortunately for Dirac, the philosopher Charlie Broad, who had studied the
-Einstein's theory, was appointed as the Professor of Philosphy at
+Fortunately for Dirac, the philosopher Charlie Broad, who had studied
+the Einstein's theory, was appointed as the Professor of Philosphy at
 Bristol in Dirac's final year, and gave lectures on relativity. From
@@
 Dirac thought that ``even approximate theories can have mathematical
-beauty'' \cite[pg. 45]{strangest_man} as a result of his engineering training,
-and this point of view of science as approximations has influenced
-Richard Feynman and many other physicists today.
+beauty'' \cite[pg. 45]{strangest_man} as a result of his engineering
+training, and this point of view of science as approximations has
+influenced many other physicists today.
 
@@
 been influenced greatly by this work, since years later he would
-merge relativity and quantum mechanics.
+merge special relativity and quantum mechanics.
 
@@
 
-What was strange to Heisenberg was that the order in which matrices are
-multiplied matters, $a$ times $b$ could be different from $b$ times $a$.
-This $ab - ba$ caught Dirac's attention, as Dirac was familiar with this
-noncommuting multiplication from his mathematics training, but did not
-understand its physical significance or implications. Always interested
-in relativity, Dirac also tried to create a relativistic version of
-Heisenberg's theory but had no success and soon gave up.
+Heisenberg's theory was strange to physicists because the order in which
+matrices are multiplied matters, $a$ times $b$ could be different from
+$b$ times $a$, so that $ab - ba$ is not zero..  This $ab - ba$ caught
+Dirac's attention, as Dirac was familiar with this noncommuting
+multiplication from his mathematics training, but did not understand its
+physical significance or implications. Always interested in relativity,
+Dirac also tried to create a relativistic version of Heisenberg's theory
+but had no success and soon gave up.
 
@@
 
+When Dirac first heard of Schrodinger's wave mechanics, he did not take
+it very seriously, as he already had a working quantum mechanics from
+Heisenberg and his own work. However, Heisenberg told Dirac to pay more
+attention to Schrodinger's version of quantum mechanics. From reading
+Schrodinger and other people's work, he realized that he could have
+gotten Schrodinger's equation from Heisenberg's theory and his own
+developments with Poisson brackets.
+
+Working with Schrodinger's wavefunctions, Dirac thought about how to
+represent two or more identical particles using wavefunctions. Dirac
+noticed that if two identical particles exchanged places, no experiment
+could possibly be done to tell that they have swapped. Following
+Heisenberg's influence of only talking about things that could be
+measured, Dirac realized that quantum theory should not predict a
+measurable change when two identical particles are swapped. Dirac worked
+out that the wavefunction must either stay the same or be multiplied by
+$-1$ when two identical particles swap in order for quantum theory to be
+consistent with this fact. Dirac found that fermions' wavefunctions
+would be multiplied by $-1$ while bosons' wavefunctions would stay the
+same. Fermions, an example being the electron, are particles that have
+the property that no two fermions can occupy the same state (the Pauli
+exclusion principle), and in some sense have a tendency to avoid each
+other.  Bosons, an example being the photon, are particles that can have
+infinitely many of them occupying the same state, and in some sense tend
+to clump.
+
+Continuing to play with the mathematics of Schrodinger's equation, Dirac
+wondered what would happen if he made the wavefunction itself one of
+Heisenberg's noncommuting quantities. From this line of thought, Dirac
+worked out a novel mathematical way to represent fermions and bosons,
+and this method is now called second quantization. Using this method, he
+was able to explain Planck's blackbody radiation formula using the new
+quantum mechanics by identifying bosons with oscillators. Second
+quantization has become one of Dirac's greatest contributions to
+physics, and it came from one of Dirac's unusual approaches to physics
+of playing with equations, looking for beautiful mathematics, and seeing
+what they predicted, as opposed to having a clear goal in mind and
+trying to develop a method for it.
+
+Dirac's successes in working with quantum mechanics must have given him
+confidence to return to his old passion, relativity. Dirac now tackled
+the problem of merging special relativity with quantum mechanics. Dirac
+again resorted to playing with mathematical quantities but this time had
+a clear goal which helped him limit the possibilities. He specified the
+characteristics the equation must have, and used those to guide him. The
+equation must agree with Einstein's special relativity, and also must
+agree with his version of quantum mechanics. Dirac also believed that
+the equation would be simple and beautiful. Through trial and error and
+months of hard work, Dirac eventually guessed the correct equation, now
+called the Dirac equation. Unlike Schrodinger's equation which only
+includes one wavefunction, Dirac's equation featured four interconnected
+wavefunctions. Additionally, Dirac's equation explained the existence of
+spin and magnetic field of the electron as a result of the merging of
+special relativity and quantum mechanics, a great triumph for
+theoretical physics.
+
+All was not well, however. Dirac's equation also predicted the existence
+of negative energy electron states, something that seemed absurd and was
+never observed in experiment before. This troubled Dirac greatly. He
+eventually postulated that all the negative energy states were filled
+somehow with undetected negative energy electrons. A vacancy in these
+negative energy states would appear as a positively charged hole, and
+only this hole could be detected in experiment. In 1932 at Caltech, Carl
+Anderson photographed the particle track of Dirac's hole, now called the
+positron. In a way, Dirac for the first time had predicted the existence
+of a new particle of nature from guessing an equation that would conform
+with previously known laws (quantum mechanics and relativity). This
+style of theoretical physics, of guessing the mathematics based on
+constraints, continued to help physicists predict the existence of more
+particles, with the Higgs boson being the most recently discovered one,
+but having been predicted by theory many years earlier.
+
+Dirac's successes in producing physical theories from playing with
+mathematics led him to believe that the ``most powerful method of
+advance...\ is to employ all the resources of pure mathematics in
+attempts to perfect and generalise the mathematical formalism that forms
+the existing basis of theoretical physics, and \emph{after} each success
+in this direction, to try to interpret the new mathematical features in
+terms of physical entities'' \cite{monopoles}.
+
+Following his own advice, Dirac generalized the mathematical form of
+Schrodinger's wavefunctions and reinterpreted his new wavefunction as
+describing particles in a magnetic field. Investigating the properties
+of these generalized wavefunctions further, Dirac discovered that if
+magnetic monopoles existed, it would explain the quantization of
+electric charge. Physicists have only found free particles or ions with
+an electric charge being an integer multiple of the charge of an
+electron, but now Dirac has a theoretical explanation for this fact.
+Unfortunately, his explanation rests on the existence of magnetic
+monopoles, and to this day no experiment has found one yet.
+
 \nocite{*}

### 2014-06-08 latex/paper2.tex
i think i'm done

@@
 
-Dirac's style and approach to his work has had an enormous influence
-on theoretical physics. To fully understand Dirac's style of thinking,
-we must begin by looking at Dirac's education and the events that
-influenced him.
+At the turn of the 20th century, many physicists shifted their focus to
+investigating the structure of matter and light, being inspired by the
+discovery of radioactivity and the electron. Around twenty years later,
+their investigations had led them to a new quantum mechanics to describe
+the behavior of the atom.  Paul Dirac was one of the most creative
+geniuses in theoretical physics during the quantum revolution. Dirac's
+style and approach to his work has had an enormous influence on
+theoretical physics. In this essay, we explore Dirac's style of physics
+by following him through his many accomplishments, but to fully
+understand Dirac's style of thinking, we must begin by looking at
+Dirac's education and the events that influenced him.
 
@@
 
+Dirac had a large influence on the physicists Schwinger and Feynman.
+Both had read Dirac's book \textit{The Principles of Quantum Mechanics}
+during their studies, and were inspired to develop a consistent quantum
+electrodynamics, which they accomplished through renormalization.
+However, Dirac, always looking for mathematical beauty in physical laws,
+felt the renormalization approach was ugly. He never accepted Schwinger
+and Feynman's quantum electrodynamics as a satisfactory theory, despite
+its successes at predicting experimental results. Dirac was never
+successful developing a quantum electrodynamics that satisfied himself
+either. In the last paragraph of his book, he writes, ``It would seem
+that we have followed as far as possible the path of logical development
+of the ideas of quantum mechanics as they are at present understood. The
+difficulties, being of a profound character, can be removed only by some
+drastic change in the foundations of the theory, probably a change as
+drastic as the passage from Bohr's orbit theory to the present quantum
+mechanics'' \cite[pg. 310]{principles}.
+
+Dirac's style of thinking has left a significant mark in theoretical
+physics. He showed the power of mathematical reasoning in discovering
+new laws of nature, although his way of doing mathematics was heavily
+influenced by his engineering background. Through guessing equations,
+playing with mathematical objects, and seeking beautiful relationships
+among these objects, Dirac made some of the greatest advances in physics
+following the formulation of quantum mechanics by Heisenberg and
+Schrodinger. His style and methods have continued to help physicists
+create new laws and predict the existence of new particles, and we see
+his influence on physics everywhere even today.
+
 \nocite{*}

### 2014-06-08 latex/paper2.tex
spell check

@@
 Fortunately for Dirac, the philosopher Charlie Broad, who had studied
-the Einstein's theory, was appointed as the Professor of Philosphy at
+the Einstein's theory, was appointed as the Professor of Philosophy at
 Bristol in Dirac's final year, and gave lectures on relativity. From
@@
 
-On one of Dirac's relaxing Sunday walks, it occured to him that the
+On one of Dirac's relaxing Sunday walks, it occurred to him that the
 noncommuting matrices in Heisenberg's theory might be related to the

### 2014-06-08 latex/paper2.tex
typo

@@
 matrices are multiplied matters, $a$ times $b$ could be different from
-$b$ times $a$, so that $ab - ba$ is not zero..  This $ab - ba$ caught
+$b$ times $a$, so that $ab - ba$ is not zero. This $ab - ba$ caught
 Dirac's attention, as Dirac was familiar with this noncommuting
